---
title: "control_self"
output: html_document
author: tthejll
---

# Setup
```{r setup}
library(tidyverse)
library(psych)
library(ggcorrplot)
library(randomForest)

```

## Load data and clean data
```{r load data}

# we load in pre-processed data
d_au_annotation <- read_csv("../data/d_au_annotation.csv") %>%
  select(-...1) %>% # remove python index column
  mutate(emo_cat2 = replace_na(emo_cat2, "NA")) %>%
  mutate_if(is.character, as.factor) %>%# make all character columns to factors
  rename_with(~ str_replace(., "_r$", ""), ends_with("_r"))
  # select(-AU45)

  # mutate(across(contains("AU"), ~(. - mean(.))/sd(.),.names = "std_{.col}")) # if we want to standardize

```

# Main expressions
```{r main expression}
d_au <- d_au_annotation %>%
  select(contains("AU"))

sevenf <- fa(d_au, nfactors = 7, rotate = "oblimin", fm = "minres") # chosent to be best fit
print(sevenf, cut = 0.3)

```


### Use of factors

```{r setup factor data}
# first we create stratou factors
f1_strat <- c("AU06", "AU07", "AU12")
f2_strat <- c("AU01", "AU02")
f3_strat = c("AU20", "AU25", "AU26")
f4_strat = c("AU14", "AU17", "AU23")
f5_strat = c("AU04", "AU07", "AU09")
f6_strat = c("AU10", "AU15", "AU17")

d_full_factors <- d_au_annotation %>%
  rowwise() %>%
  mutate(f1_mean_strat = ((AU06*0.982) * (AU07*0.446) * (AU12*0.882))/3,
         f4_mean_strat = ((AU14*0.724) * (AU17*0.632) * (AU23*0.710))/3,
         f4_mean = ((AU12*0.776) * (AU14*0.458))/2,
         f6_mean = AU06*0.817) %>%
  ungroup() %>%
  select(contains("AU"), contains("mean"), control_self, ppt) %>%
  filter(!is.na(control_self))

```


```{r factor 1}

# choose the right factor
d_temp_au <- d_full_factors %>%
  select(contains("f1"), control_self, ppt) %>%
  filter(!is.na(control_self)) %>%
  mutate(frame_n = 1:nrow(.),
         indicator = if_else(ppt != lag(ppt, default = first(ppt)), 1, 0)) %>%
  group_by(ppt) %>%
  mutate(frames = 1:n()) %>% ungroup()

# Calculate mean for each participant and condition
condition_means <- d_temp_au %>%
  group_by(control_self, ppt) %>%
  summarise(mean_value = mean(f1_mean_strat), sd = sd(f1_mean_strat), n = n(), .groups = "drop")

# calculate grand means
grand_mean <- d_temp_au %>%
  group_by(control_self) %>%
  summarise(mean_value = mean(f1_mean_strat),sd = sd(f1_mean_strat), n = n(), .groups = "drop")

count_zeros <- d_temp_au %>%
  filter(f1_mean_strat == 0) %>%
  group_by(control_self) %>%
  summarise(mean_value = mean(f1_mean_strat),sd = sd(f1_mean_strat), n = n(), .groups = "drop") %>%
  mutate(percent = (n/grand_mean$n)*100)

# order High, Low, Medium
percent_zeros <- count_zeros$n/grand_mean$n*100


# labelling help
ppt_levels <- c("301_neg2", "305_neg2", "306_neg1", "308_neg1", "309_neg1", "309_neg2", "312_neg1", "313_neg2", "316_neg2", "318_neg2", "319_neg1", "323_neg2", "324_neg2", "326_neg2", "327_neg1", "328_neg2", "332_neg1", "333_neg2", "334_neg1", "335_neg2", "338_neg2", "339_neg1", "341_neg2", "391_neg1")

ppt_labels <- c(1:4, "5 - neg1", "5 - neg2", 6:23)

ppt_label_map <- setNames(ppt_labels, ppt_levels)


# plot overall effect  - saving ratio 800x450
d_temp_au %>%
  ggplot(
    data = .,
    aes(x = frames,
             y = f1_mean_strat,
             group = 1,
             )) +
  geom_line(size = 0.2, alpha = 0.07) +
  geom_hline(
    data = grand_mean,
    aes(yintercept = mean_value, color = control_self), lty = "dashed") +
  scale_color_viridis_d(name="Control-self") + 
  theme_minimal() + 
  ylim(c(0,1)) +
  ylab("Enjoyment smile activation") + 
  xlab("Frames") +
  labs(title = "Weighted average activation of the Enjoyment Smile AUs")


# plot participant effect - saving ratio 800x550
d_temp_au %>%
  ggplot(
    data = .,
    aes(x = frames,
             y = f1_mean_strat,
             group = 1,
             )) +
  geom_line(size = 0.2, alpha = 0.5) +
  facet_wrap(~ppt, labeller = labeller(ppt = ppt_label_map), nrow = 6, ncol = 4) +
  geom_hline(
    data = condition_means,
    aes(yintercept = mean_value, color = control_self), lty = "dashed") +
  scale_color_viridis_d(name="Control-self") + 
  theme_minimal() +
  ylim(c(0,1)) +
  ylab("Enjoyment smile activation") + 
  xlab("Frames") + 
  labs(title = "Weighted average activation of the Enjoyment Smile AUs (per participant)")

# calculate overall effects
# l v H
d_temp_au %>%
  filter(control_self != "Medium") %>%
  wilcox.test(f1_mean_strat ~ control_self, data = .)

# L v M
d_temp_au %>%
  filter(control_self != "High") %>%
  wilcox.test(f1_mean_strat ~ control_self, data = .)

# High v Medium
d_temp_au %>%
  filter(control_self != "Low") %>%
  wilcox.test(f1_mean_strat ~ control_self, data = .)



```

```{r f1 per participant analysis}
condition_means %>% 
  ggplot(data = .,
         aes(x = mean_value,
             color = control_self,
             fill = control_self)) + 
  geom_density(alpha = 0.1) + 
  scale_fill_viridis_d(name="Control-self") +
  scale_color_viridis_d(name="Control-self") + 
  theme_minimal() + 
  labs(title = "Density plot for participant mean activation of Enjoyment Smile",
       x = "Mean activation",
       y = "Desnsity")

# print means for participants
condition_means %>%
  filter(ppt != "301_neg2") %>%
  group_by(control_self) %>%
  summarise(mean = mean(mean_value), SD = sd(mean_value), n = n())

# low v high
condition_means %>%
  filter(control_self != "Medium") %>%
  wilcox.test(mean_value ~ control_self, data = .)

# low v medium
condition_means %>%
  filter(control_self != "High") %>%
  wilcox.test(mean_value ~ control_self, data = .)

# medium v high
condition_means %>%
  filter(control_self != "Low") %>%
  wilcox.test(mean_value ~ control_self, data = .)


```

# feature importance

```{r feature importance from python}

# copied from python - control-self
feature_name <- c('f7_sd', 'f2_sum', 'f2_max', 'f7_max', 'f2_sd', 'f2_mean', 'f7_mean', 'f7_sum', 'f3_max', 'f3_mean', 'f3_sd', 'f3_sum', 'f6_mean', 'f5_sd', 'f1_max', 'f4_sd', 'f1_sd', 'f4_max', 'f1_sum', 'f1_mean', 'f4_mean', 'f4_sum', 'f5_mean', 'f5_sum', 'f5_max')
feature_importance_val <- c(0.020280697504879514, 0.025190966461652496, 0.025697583687637936, 0.02570409569845856, 0.025836824501957522, 0.02824834586818671, 0.02834762331992622, 0.028398165444305228, 0.028420530859477613, 0.029601804948587728, 0.03042343448074982, 0.0319342782317302, 0.037426871809268576, 0.03979312629485129, 0.04092628791365906, 0.041970709471457444, 0.043525095263642684, 0.0471470734235623, 0.05030419379230158, 0.053179616786487235, 0.05558970331981823, 0.05725266959537617, 0.059872811469457485, 0.06982477232229535, 0.07510271753027309) %>% round(digits = 4)
feature_importance <- tibble(feature_name, feature_importance_val) %>%
  mutate(feature_name = as.factor(feature_name),
         appraisal = "Control-self") %>%
  separate(feature_name, c("Factor", "val"), sep = "_", remove = FALSE)
  

feature_importance %>%
ggplot(data = .,
       aes(
         x = fct_reorder(feature_name, feature_importance_val), 
         y = feature_importance_val)) +
  coord_flip() +
  geom_col(aes(fill = Factor)) + 
  scale_fill_viridis_d() +
  theme_minimal() +
  ggtitle("Feature importance in RandomForest Classifier") +
  ylab("Decrease in impurity") +
  xlab("")

```

```{r feature importance AU from python}

# copied from python - control-self
feature_name <- c('AU05', 'AU02', 'AU20', 'AU23', 'AU09', 'AU45', 'AU01', 'AU15', 'AU07', 'AU17', 'AU14', 'AU10', 'AU04', 'AU06', 'AU12', 'AU26', 'AU25')
feature_importance_val <- c(0.016490137145388488, 0.02805084570725775, 0.03907909202320244, 0.040411701145889115, 0.04203659350461894, 0.04250977647921853, 0.04337311211802373, 0.044447042305981094, 0.0551990295433165, 0.066089173307346, 0.07114878246848244, 0.07190293959149537, 0.07420195577790754, 0.0751708215271029, 0.0822206940854503, 0.09043568069716515, 0.11723262257215376) %>% round(digits = 4)
feature_importance <- tibble(feature_name, feature_importance_val) %>%
  mutate(feature_name = as.factor(feature_name), 
         appraisal = "Control-self")
  

feature_importance %>%
ggplot(data = .,
       aes(
         x = fct_reorder(feature_name, feature_importance_val), 
         y = feature_importance_val)) +
  coord_flip() +
  geom_col() + 
  scale_fill_viridis_d() +
  theme_minimal() +
  ggtitle("Feature importance in RandomForest Classifier") +
  ylab("Decrease in impurity") +
  xlab("")

```
```{r feature importance Stratou from python}

f1_strat <- c("AU06", "AU07", "AU12")
f2_strat <- c("AU01", "AU02")
f3_strat = c("AU20", "AU25", "AU26")
f4_strat = c("AU14", "AU17", "AU23")
f5_strat = c("AU04", "AU07", "AU09")
f6_strat = c("AU10", "AU15", "AU17")


# copied from python - self
feature_name <- c('f2_sd', 'f2_max', 'f2_mean', 'f2_sum', 'f6_sd', 'f5_sd', 'f5_max', 'f4_max', 'f6_max', 'f5_sum', 'f6_mean', 'f5_mean', 'f6_sum', 'f4_sum', 'f4_mean', 'f1_sd', 'f1_max', 'f4_sd', 'f1_sum', 'f1_mean', 'f3_sum', 'f3_mean', 'f3_sd', 'f3_max')
feature_importance_val <- c(0.019021744610932367, 0.01979708956863626, 0.021775914432399436, 0.023701209590687475, 0.03603393284995201, 0.03863699089568017, 0.03865013423935222, 0.038997661088222675, 0.03978659985868629, 0.03984979127131145, 0.03997715719508615, 0.04013369094334502, 0.04085106547641898, 0.04188244736687314, 0.04349990909106514, 0.043721304146244595, 0.045335932563415116, 0.04645294512142578, 0.05088416116824198, 0.05287992324706663, 0.05490599913653369, 0.05765263364933102, 0.061258446420389356, 0.06431331606870302) %>% round(digits = 4)
feature_importance <- tibble(feature_name, feature_importance_val) %>%
  mutate(feature_name = as.factor(feature_name),
         appraisal = "Control-self") %>%
  separate(feature_name, c("Factor", "val"), sep = "_", remove = FALSE)
  

feature_importance %>%
ggplot(data = .,
       aes(
         x = fct_reorder(feature_name, feature_importance_val), 
         y = feature_importance_val)) +
  coord_flip() +
  geom_col(aes(fill = Factor)) + 
  scale_fill_viridis_d() +
  theme_minimal() +
  ggtitle("Feature importance in RandomForest Classifier") +
  ylab("Decrease in impurity") +
  xlab("")

```

#### Factor analysis for control-self - this isn't used in the analyss.

It was tested to see if there were differences in how expressions were used when looking at different levels of appraisal

# Model control-self

## control-self high
```{r filter control self}
# d_au_self_high <- d_au_annotation %>%
#   filter(control_self == "High") %>%
#   select(contains("AU"))
# 
# describe(d_au_self_high)

```

### data suitability
``` {r data checks}
# Kaiser-Meyer-Olkin factor adequacy
# d_au_self_high %>%
#   KMO(.) # fine 
# 
# d_au_self_high %>%
#   cor(.) %>%
# cortest.bartlett(., n = nrow(d_au_self_high))

```

### Number of factors to retain
```{r Minimum Average Partial correlation} 
# 
# d_au_self_high %>%
# VSS(., use ="complete.obs", fm = 'mle', rotate = "oblimin") %>% print()
# 
# d_au_self_high %>%
#   fa.parallel(., main = "Parallel Analysis", n.iter = 100)
# 
# 
# d_au_self_high %>%
#   scree(factors = F,)
```

Based on MAP and parallel analysis, we want to test everything from 1 to 7 factors

### Testing different factors

```{r factor analysis}

# onef <- fa(d_au_self_high, nfactors = 1, rotate = "oblimin", fm = "minres")
# print(onef, cutoff = 0.3)
# 
# twof <- fa(d_au_self_high, nfactors = 2, rotate = "oblimin", fm = "minres")
# print(twof, cutoff = 0.3)
# 
# threef <- fa(d_au_self_high, nfactors = 3, rotate = "oblimin", fm = "minres")
# print(threef, cutoff = 0.3)
# 
# fourf <- fa(d_au_self_high, nfactors = 4, rotate = "oblimin", fm = "minres")
# print(fourf, cutoff = 0.3)
# 
# fivef <- fa(d_au_self_high, nfactors = 5, rotate = "oblimin", fm = "minres")
# print(fivef, cutoff = 0.3)
# 
# sixf <- fa(d_au_self_high, nfactors = 6, rotate = "oblimin", fm = "minres") # has ultra heywood case
# print(sixf, cutoff = 0.3)
# 
# sevenf <- fa(d_au_self_high, nfactors = 7, rotate = "oblimin", fm = "minres") # chosent to be best fit
# print(sevenf, cutoff = 0.3)
# 
# eigthf <- fa(d_au_self_high, nfactors = 8, rotate = "oblimin", fm = "minres", max.iter = 1000)
# print(eigthf, cutoff = 0.3)
# 
# scree(d_au_self_high)

```

## control-self medium #######################################################
```{r filter control self}
# d_au_self_medium <- d_au_annotation %>%
#   filter(control_self == "Medium") %>%
#   select(contains("AU"))

```

### data suitability
``` {r data checks}
# # Kaiser-Meyer-Olkin factor adequacy
# d_au_self_medium %>%
#   KMO(.) # fine 
# 
# d_au_self_medium %>%
#   cor(.) %>%
# cortest.bartlett(., n = nrow(d_au_self_medium))

```

### Number of factors to retain
```{r Minimum Average Partial correlation} 

# d_au_self_medium %>%
# VSS(., use ="complete.obs", fm = 'mle', rotate = "oblimin") %>% print()
# 
# d_au_self_medium %>%
#   fa.parallel(., main = "Parallel Analysis", n.iter = 100)
# 
# 
# d_au_self_medium %>%
#   scree(factors = F,)
```

Based on MAP and parallel analysis, we want to test everything from 1 to 7 factors

### Testing different factors

```{r factor analysis}
# 
# onef <- fa(d_au_self_medium, nfactors = 1, rotate = "oblimin", fm = "minres")
# print(onef, cutoff = 0.3)
# 
# twof <- fa(d_au_self_medium, nfactors = 2, rotate = "oblimin", fm = "minres")
# print(twof, cutoff = 0.3)
# 
# threef <- fa(d_au_self_medium, nfactors = 3, rotate = "oblimin", fm = "minres")
# print(threef, cutoff = 0.3)
# 
# fourf <- fa(d_au_self_medium, nfactors = 4, rotate = "oblimin", fm = "minres")
# print(fourf, cutoff = 0.3)
# 
# fivef <- fa(d_au_self_medium, nfactors = 5, rotate = "oblimin", fm = "minres")
# print(fivef, cutoff = 0.3)
# 
# sixf <- fa(d_au_self_medium, nfactors = 6, rotate = "oblimin", fm = "minres") # has ultra heywood case
# print(sixf, cutoff = 0.3)
# 
# sevenf <- fa(d_au_self_medium, nfactors = 7, rotate = "oblimin", fm = "minres") # chosent to be best fit
# print(sevenf, cutoff = 0.3)
# 
# eigthf <- fa(d_au_self_medium, nfactors = 8, rotate = "oblimin", fm = "minres", max.iter = 1000)
# print(eigthf, cutoff = 0.3)
# 
# scree(d_au_self_medium)

```

## control-self low #######################################################
```{r filter control self}
# d_au_self_low <- d_au_annotation %>%
#   filter(control_self == "Low") %>%
#   select(contains("AU"))

```

### data suitability
``` {r data checks}
# Kaiser-Meyer-Olkin factor adequacy
# d_au_self_low %>%
#   KMO(.) # fine 
# 
# d_au_self_low %>%
#   cor(.) %>%
# cortest.bartlett(., n = nrow(d_au_self_low))

```

### Number of factors to retain
```{r Minimum Average Partial correlation} 

# d_au_self_low %>%
# VSS(., use ="complete.obs", fm = 'mle', rotate = "oblimin") %>% print()
# 
# d_au_self_low %>%
#   fa.parallel(., main = "Parallel Analysis", n.iter = 100)
# 
# 
# d_au_self_low %>%
#   scree(factors = F,)
```

Based on MAP and parallel analysis, we want to test everything from 1 to 7 factors

### Testing different factors

```{r factor analysis}

# onef <- fa(d_au_self_low, nfactors = 1, rotate = "oblimin", fm = "minres")
# print(onef, cutoff = 0.3)
# 
# twof <- fa(d_au_self_low, nfactors = 2, rotate = "oblimin", fm = "minres")
# print(twof, cutoff = 0.3)
# 
# threef <- fa(d_au_self_low, nfactors = 3, rotate = "oblimin", fm = "minres")
# print(threef, cutoff = 0.3)
# 
# fourf <- fa(d_au_self_low, nfactors = 4, rotate = "oblimin", fm = "minres")
# print(fourf, cutoff = 0.3)
# 
# fivef <- fa(d_au_self_low, nfactors = 5, rotate = "oblimin", fm = "minres")
# print(fivef, cutoff = 0.3)
# 
# sixf <- fa(d_au_self_low, nfactors = 6, rotate = "oblimin", fm = "minres") # has ultra heywood case
# print(sixf, cutoff = 0.3)
# 
# sevenf <- fa(d_au_self_low, nfactors = 7, rotate = "oblimin", fm = "minres") # chosent to be best fit
# print(sevenf, cutoff = 0.3)
# 
# eigthf <- fa(d_au_self_low, nfactors = 8, rotate = "oblimin", fm = "minres", max.iter = 1000)
# print(eigthf, cutoff = 0.3)
# 
# scree(d_au_self_low)

```
